---
title: "Testing Out Sparklyr on DNAnexus"
author: "Ted Laderas"
date: "6/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sparklyr)
library(tidyverse)
sc <- spark_connect(master="local")
```

This notebook mostly follows everything that is available here, with my learnings alongside it.

```{r}
library(dplyr)
iris_tbl <- copy_to(sc, iris)
flights_tbl <- copy_to(sc, nycflights13::flights, "flights")
batting_tbl <- copy_to(sc, Lahman::Batting, "batting")
dplyr::src_tbls(sc)
```

```{r}
colnames(flights_tbl)
```

## Distribution of arrival delays

Using the conventional ggplot2 functions: 

```{r}
flights_tbl %>%
  ggplot2::ggplot() +
  aes(x=arr_delay) +
  geom_histogram()
```


```{r}
sparklyr::
```


```{r}
flights_tbl %>%
  filter(year > 1990) %>%
  mutate(arr_delay
```

```{r}
partitions <- flights_tbl %>%
  filter(year > 2000) %>%
  mutate(time_hour = as.numeric(time_hour),
         arr_delay = as.numeric(arr_delay)
         ) %>%
  sdf_partition(training = 0.8, test = 0.2, seed = 1099)

partitions$training
```

```{r}
# fit a linear model to the training dataset`
fit <- partitions$training %>%
  ml_linear_regression(response = "arr_delay", features = c("time_hour"))
fit
```

## Machine Learning with Spark

sparklyr provides three families of functions that you can use with Spark machine learning:

- Machine learning algorithms for analyzing data (ml_*)
- Feature transformers for manipulating individual features (ft_*)
- Functions for manipulating Spark DataFrames (sdf_*)

## MLib Workflow

1. Perform SQL queries through the sparklyr dplyr interface,
2. Use the sdf_* and ft_* family of functions to generate new columns, or partition your data set,
3. Choose an appropriate machine learning algorithm from the ml_* family of functions to model your data,
4. Inspect the quality of your model fit, and use it to make predictions with new data.
5. Collect the results for visualization and further analysis in R

